% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  11pt,
  a4paper,
  twoside,
  openright]{scrbook}
\usepackage{xcolor}
\usepackage[top=25mm,bottom=25mm,left=30mm,right=25mm]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{5}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother


\usepackage{longtable,booktabs,array}
\newcounter{none} % for unnumbered tables
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother





\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}



 


\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{bookmark}{}{\usepackage{bookmark}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={A Decision-Centred Modelling Environment for Long-Horizon Planning Under Deep Uncertainty},
  pdfauthor={Seyed Ahmad Mahmoudi Lahijani},
  colorlinks=true,
  linkcolor={0.2,0.4,0.6},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}


\title{A Decision-Centred Modelling Environment for Long-Horizon
Planning Under Deep Uncertainty}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Foundations, Architecture, and Application to Industrial
Process Heat Decarbonisation}
\author{Seyed Ahmad Mahmoudi Lahijani}
\date{2025}
\begin{document}
\frontmatter
\maketitle

\renewcommand*\contentsname{Table of contents}
{
\setcounter{tocdepth}{2}
\tableofcontents
}

\mainmatter
\bookmarksetup{startatroot}

\chapter{A Decision-Centred Modelling
Environment}\label{a-decision-centred-modelling-environment}

\bookmarksetup{startatroot}

\chapter*{Welcome}\label{welcome}
\addcontentsline{toc}{chapter}{Welcome}

\markboth{Welcome}{Welcome}

This is the online companion to the manuscript:

\textbf{A Decision-Centred Modelling Environment for Long-Horizon
Planning Under Deep Uncertainty: Foundations, Architecture, and
Application to Industrial Process Heat Decarbonisation}

\emph{Seyed Ahmad Mahmoudi Lahijani}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section*{Quick Navigation}\label{quick-navigation}
\addcontentsline{toc}{section}{Quick Navigation}

\markright{Quick Navigation}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
If you want to\ldots{}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Go to\ldots{}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Understand the framework's big picture &
\href{front-matter/context-declaration.qmd}{Context Declaration} \\
Know how to read this document &
\href{front-matter/navigation-guide.qmd}{Navigation Guide} \\
Start the conceptual argument &
\href{manuscript/module-0-orientation.qmd}{Module 0: Orientation} \\
See what has been built & \href{manuscript/module-6-poc.qmd}{Module 6:
The Proof of Concept} \\
Ask the framework a question & \hyperref[ai-panel]{AI Discussion
Panel} \\
Contribute an extension & \href{../CONTRIBUTING.md}{Contributing} \\
\end{longtable}
}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section*{The Core Finding}\label{the-core-finding}
\addcontentsline{toc}{section}{The Core Finding}

\markright{The Core Finding}

\begin{quote}
When the Edendale electrification pathway is evaluated against a
structured ensemble of 64 futures, \textbf{23 of those futures} produce
grid hosting capacity exceedances that generate regional infrastructure
costs invisible from within the site boundary alone.

This result --- 36 percent of the ensemble revealing a constraint that
site-level analysis cannot see --- is the gap this framework fills, made
analytically visible.
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section*{The Central Commitment}\label{the-central-commitment}
\addcontentsline{toc}{section}{The Central Commitment}

\markright{The Central Commitment}

The framework is organised around a single foundational reversal:

\begin{quote}
\emph{The decision makes the boundary; the artefact makes the
connection.}
\end{quote}

This document is a living resource. Contributions, extensions, and
critiques are welcomed. See the \hyperref[community-extension]{Community
Extension} section of the Context Declaration.

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, bottomrule=.15mm, bottomtitle=1mm, breakable, colback=white, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, coltitle=black, left=2mm, leftrule=.75mm, opacityback=0, opacitybacktitle=0.6, rightrule=.15mm, title=\textcolor{quarto-callout-note-color}{\faInfo}\hspace{0.5em}{AI-Assisted Engagement}, titlerule=0mm, toprule=.15mm, toptitle=1mm]

Upload this document to your preferred AI assistant to ask
framework-specific questions. The Context Declaration is written as a
system prompt for AI agents. The Node Declaration at the start of every
section tells AI readers what they can skip and what they must read.

\end{tcolorbox}

\part{Front Matter}

\chapter{Context Declaration}\label{context-declaration}

For Human Readers, AI Assistants, and the Research Community

\hfill\break

\section{A Vision Before the Detail}\label{a-vision-before-the-detail}

This document describes a framework for making better decisions under
deep uncertainty. Before entering the architecture, the vocabulary, or
the proof of concept, it is worth pausing on what that means in
practice, and what it could mean at scale.

Consider a large dairy processor evaluating whether to electrify its
heat supply by 2035. The decision looks, on the surface, like a
site-level engineering and finance problem: compare the capital cost of
electric boilers against biomass conversion, estimate operating costs,
account for the carbon price, choose the pathway with the lower net
present value. That framing is not wrong. It is incomplete. The pathway
choice propagates consequences to a regional electricity grid that may
not have the capacity to absorb the additional load without
reinforcement. That reinforcement, if required, takes years to plan,
consent, and build. It also draws on the same grid capacity that other
industrial users in the region are simultaneously seeking to access. The
biomass pathway, meanwhile, draws on a regional feedstock supply that
other users are simultaneously competing for. Both pathways are exposed
to a carbon price whose trajectory is set by a policy process that is
institutionally contested and whose design has changed materially
several times in the past decade. And the facility's decision will be
made by a corporate sustainability team, reviewed by a board with both
financial and reputational objectives, influenced by a government
co-investment programme with its own eligibility criteria and funding
cycles, shaped by community expectations about the future of the
Southland economy, and ultimately constrained by physical infrastructure
that was built for a different industrial era.

None of this complexity is unusual. It is the normal texture of a
consequential infrastructure decision in the early twenty-first century.
The problem is that the analytical tools most commonly brought to bear
on this kind of decision were designed for a simpler version of it: one
site, one future, one objective, one model boundary. They are not
inadequate because they are unsophisticated. They are inadequate because
they are organised around the wrong centre of gravity. They are
organised around the system rather than around the decision.

This framework is organised around the decision. That reversal, simple
to state and demanding to operationalise, is what the rest of this
document develops.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{What the Framework Is
Attempting}\label{what-the-framework-is-attempting}

The framework developed here begins from a proposition that is easy to
accept in the abstract and surprisingly difficult to implement in
practice: the model should develop toward the parameters that influence
the decision, not toward the completeness of the system description.

In conventional modelling practice, a system is first described as
completely as possible within a boundary defined by physical or
technical extent, and the decision is then framed as a query to that
model. The model defines what can be asked. In the framework proposed
here, the process is reversed. The decision is articulated first: what
is being chosen, by whom, with consequences at which scales, and under
which dimensions of uncertainty. The model then develops toward the
parameters that influence that decision. Physical detail is incorporated
where it changes the outcome of the comparison. Institutional conditions
are represented where they shift the ranking of alternatives across
futures. Social and environmental consequences are included where they
are relevant to what the decision-maker must account for, not as
afterthoughts appended to a completed technical analysis but as
first-class dimensions of the decision frame from the outset.

This means the framework grows. It begins with a minimum viable
representation that is honest about its simplifications and explicit
about where additional detail would change the decision-relevant
comparison. At each development step, it asks: what is the most
influential parameter not yet well represented? What constraint, once
modelled, would shift the robustness ranking of the alternatives? What
future condition, once included, would reveal a vulnerability currently
invisible? The answer to these questions, rather than a commitment to
physical completeness, determines what gets modelled next.

The consequence is an analytical environment that is at once humble and
ambitious. Humble, because it never claims to have modelled everything
that matters. Ambitious, because it claims to have modelled exactly what
matters most at each stage of development, and to be able to demonstrate
that claim through the traceability of its artefacts and the
transparency of its decision-relevant comparisons.

There is a further property that makes this ambition practically
important. The framework is self-directing in its development. At each
stage, the regret sensitivity diagnostics and the scenario discovery
analysis of the DMDU orchestration layer identify which uncertain
drivers most strongly determine pathway preference and therefore which
simplification, once removed, would most change the decision-relevant
comparison. The framework lights up the domain that needs to be explored
next. It does not promise to eliminate uncertainty. It promises to make
decisions more intelligible in its presence, and to reveal, at each
step, where the analytical torch should be turned next.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{The Broader Vision: What Full Deployment Could
Achieve}\label{the-broader-vision-what-full-deployment-could-achieve}

The Edendale--Southland proof of concept documented in this manuscript
is, by the framework's own declared standards, a first-generation
demonstration. It demonstrates that the architecture works, that the
site-to-region coupling reveals consequences invisible from within the
site boundary, and that the artefact governance chain makes those
findings traceable and reproducible. But the Edendale case is one node
in a far larger decision landscape, and it is worth being explicit about
what the framework could achieve if deployed at greater scale and depth.

At the corporate scale, a single company such as Fonterra operates
multiple large processing sites across New Zealand, each with its own
pathway decision, each connected to regional grid infrastructure at
different GXPs, each drawing on regional biomass and other resource
systems with different supply chain characteristics. The pathway
decision at Edendale does not exist in isolation from the decisions
being made at Canpac, Waitoa, Darfield, or Clandeboye. These decisions
interact through shared grid infrastructure, shared biomass markets, and
a shared carbon pricing environment. A company-scale deployment of the
framework would evaluate all site-level pathway decisions
simultaneously, using a shared regional electricity module that
represents the aggregate demand pressure on the national transmission
system, a shared biomass resource module that represents competition for
regional feedstock supply, and a shared policy module that represents
the ETS carbon price trajectory and its interaction with the investment
timing of all facilities.

Beyond strategic pathway planning, the same architecture supports
operational optimisation. Once a pathway has been selected and
implemented, the site module can be configured to support day-ahead and
week-ahead dispatch optimisation, using real-time operational data from
the facility's control systems translated through the artefact
interface. Existing energy management system data, existing engineering
assessments, and existing financial models all translate into governed
artefacts without requiring the organisation to start from scratch. The
strategic and operational layers of the same organisation operate within
the same analytical environment, connected through the same backbone:
strategic pathway analysis informs which assets are available;
operational dispatch analysis demonstrates how those assets perform day
to day; and the accumulated operational record informs the next round of
strategic reassessment.

The result of a corporate deployment would not be an optimal transition
plan, because no such plan can be specified under the conditions of deep
uncertainty the framework acknowledges. It would instead be a robust
transition strategy: a set of sequencing and staging decisions that
remains defensible across many plausible futures, where those futures
include not only grid and biomass supply conditions but also the
decisions of competing users, the responses of regulators, and the
evolution of the corporate sustainability reporting environment.

At the sector and regional scale, the same architecture extends to
encompass multiple companies and multiple sectors. The Southland
region's dairy, meat, and wood products processors share the same
regional grid, the same biomass resource zone, and the same policy
environment. A regional deployment of the framework would represent all
major industrial heat users simultaneously, allowing the analytical
environment to evaluate not only individual site decisions but the
system-level consequences of different regional transition
configurations. This is a question that no individual site assessment
can answer, that no national model currently addresses at the required
spatial and temporal resolution, and that the framework's architecture
is specifically designed to support.

At the national scale, the framework's architecture is compatible with
coupling to the national-scale models that already exist. TIMES-NZ, New
Zealand's national energy system model, produces economy-wide scenario
outputs including electricity price trajectories, sectoral fuel demand
projections, and national emissions pathways. These outputs represent
exactly the kind of exogenous future conditions that belong in a
FutureArtefact ensemble. A national-scale deployment would use TIMES-NZ
scenario outputs to calibrate the ranges of the uncertain drivers in the
ensemble, ensuring that site-level and regional-level pathway
comparisons are consistent with plausible national energy trajectories.
The initial coupling would be one-directional: TIMES-NZ scenarios inform
the ensemble design. The eventual next phase would implement the reverse
coupling, allowing the aggregate industrial pathway decisions revealed
by the framework to be summarised as demand scenarios and passed back to
TIMES-NZ, so that national and regional levels co-evolve iteratively
through the Gauss-Seidel coupling architecture described in the
framework's next-phase specification.

At the domain level, the framework's decision-first architecture is not
confined to energy. Water system planners in a drought-exposed
catchment, regional transport authorities evaluating infrastructure
investment under uncertain demand, land-use planners navigating
competing development pressures, and industrial decarbonisation
investors in Southeast Asia, the European industrial heartland, and the
Gulf petrochemical complex all face the same structural problem:
long-lived commitments, interacting infrastructure layers, deep
uncertainty about future conditions, and the need for decisions that are
accountable to multiple stakeholders with different perspectives. Each
of these domains can be instantiated within the framework's architecture
by specifying the relevant decision class, the relevant module
structure, and the relevant artefact families. The thin-waist principle
ensures that contributions from specialists in each domain can be
integrated without requiring those specialists to understand or maintain
the full framework.

At the stakeholder accountability level, the framework's artefact
governance architecture creates something genuinely new in planning
practice: a traceable record of how a decision-relevant finding was
produced, from the raw input data and configuration parameters through
the module runs and acceptance gates to the regret and robustness
metrics that informed the decision. This traceability is not merely a
technical property. It is an institutional one. A decision made on the
basis of framework analysis can be audited: a regulator, a community
representative, or a future analyst can follow the lineage from the
finding back to the assumptions and ask, precisely and specifically,
whether the assumptions were appropriate. It becomes possible to say,
with specificity: under these conditions about grid headroom, biomass
supply, and carbon price trajectory, the electrification pathway
generates lower regret than the biomass pathway in 41 of 64 plausible
futures, and here is the complete audit trail of the analysis that
produced that finding. That is a qualitatively different foundation for
public accountability than the outputs of a black-box optimisation.

This accountability dimension extends, ultimately, to questions of
energy justice and democratic governance. When long-horizon
infrastructure decisions are made on the basis of analyses that are
traceable, multi-perspective, and open to community interrogation, the
governance of those decisions becomes more democratic in a substantive
rather than procedural sense. Communities can participate not merely by
attending consultation meetings but by engaging with the analytical
basis of the decisions that will shape their regional economies for
decades. The framework's architecture makes that engagement possible in
a governed, productive, and cumulatively valuable way.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{The Architecture That Makes This
Possible}\label{the-architecture-that-makes-this-possible}

None of the above requires building a model of everything. It requires
building an architecture that can accommodate heterogeneous analytical
components, connect them through governed interfaces, evaluate their
combined outputs under structured uncertainty, and grow progressively
toward the parameters that most influence the decisions at hand.

The thin-waist artefact structure is the mechanism that makes this
possible. It does not require that all components be simple. The site
dispatch module can be a rule-based proportional allocator in the first
generation, an LP-based scheduler in the second, and an OpenModelica
thermal network simulation in the third. The regional electricity module
can be a stylised screening model in the first generation, a full PyPSA
network optimisation in the second, and a trained ML surrogate of that
optimisation in the third. The biomass resource module can be a scalar
multiplier in the first generation and a spatially explicit logistics
model calibrated to satellite-derived forest inventory data in the
second. The macroeconomic context module can be a TIMES-NZ scenario
parameter table in the first generation and a coupled bi-directional
model exchange in a later generation. At every generation, what crosses
the module boundaries is the same: a schema-conforming artefact with
declared provenance and a validated integrity record. The rest of the
framework need not know which generation of implementation produced it.

The decision-first principle governs the order in which these
developments are pursued. The framework does not develop toward greater
complexity in general. It develops toward greater decision relevance
specifically. The question asked at each development step is: what is
the most consequential simplification currently in the model, in the
sense that removing it would most change the decision-relevant
comparison? The answer to that question, produced by the regret
sensitivity diagnostics and the scenario discovery analysis of the DMDU
orchestration layer, is the framework's own contribution to determining
what should be developed next.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Who This Is For}\label{who-this-is-for}

The framework described in this document is addressed to at least six
audiences, and it asks different things from each. What it offers all of
them is a version of the same thing: a way of organising analytical
complexity that keeps the decision visible, keeps the assumptions
traceable, and keeps the path toward better analysis open as knowledge
grows.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{It asks researchers to engage with a methodological argument
that cuts across several established traditions without belonging
entirely to any of
them.}\label{it-asks-researchers-to-engage-with-a-methodological-argument-that-cuts-across-several-established-traditions-without-belonging-entirely-to-any-of-them.}

In complex planning environments, the quality of decision support
depends more on analytical architecture than on model sophistication. A
more powerful solver embedded in a poorly framed decision environment
produces better answers to the wrong question. A well-governed,
decision-centred architecture produces answers that are less precise in
a narrow mathematical sense but more defensible in the sense that
actually matters: they are traceable to their assumptions, comparable
across alternative futures, and revisable as knowledge improves.

This is a substantive intellectual contribution to the DMDU tradition
rather than a repackaging of existing ideas. The RDM community,
represented in the work of Lempert, Popper, Bankes, Walker, Marchau,
Kwakkel, and Haasnoot, has established the evaluative standards of
regret, robustness, and satisficing and has developed the computational
methods of scenario discovery and exploratory modelling. What it has not
systematically addressed is the architectural question of how modular,
multi-scale analytical environments should be organised so that
DMDU-style exploration can operate across institutional and physical
boundaries simultaneously, with governed artefact exchange that
preserves traceability across all scales. The thin-waist artefact
architecture, the decision-first boundary principle, and the
progressive-refinement philosophy are contributions to that open
question.

Researchers working in adjacent traditions will find points of
engagement throughout. Those working in integrated assessment modelling
will find the framework's treatment of scale heterogeneity and the
site-to-region coupling directly relevant to the persistent challenge of
connecting bottom-up engineering detail to top-down economic consistency
without sacrificing either. Those working in socio-technical systems
theory will find the framework's treatment of multi-actor perspectives
and system-level regret directly relevant to the challenge of
representing institutional complexity in planning models in a form that
is analytically tractable rather than merely acknowledged. Those working
in the philosophy of science will find the framework's post-positivist
epistemological commitments, developed in Appendix I-A, a substantive
engagement with questions about what models are for and what counts as
analytical legitimacy under deep uncertainty.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{It asks energy system modellers to consider how the tools
they already use and trust can contribute to something larger than any
one of them can achieve
alone.}\label{it-asks-energy-system-modellers-to-consider-how-the-tools-they-already-use-and-trust-can-contribute-to-something-larger-than-any-one-of-them-can-achieve-alone.}

The open-source energy modelling ecosystem is mature and growing. PyPSA
provides network-aware electricity system optimisation with an active
international development community and a growing library of regional
and national model implementations. Linopy, the algebraic modelling
layer underlying PyPSA's latest architecture, makes large network
problems tractable with a clean Python interface. OpenModelica provides
equation-based physical simulation of thermal, electrical, and fluid
systems with modular, reusable component libraries suited to site-level
industrial process modelling. Calliope and OSeMOSYS provide flexible,
transparent energy system frameworks with strong open-data commitments
and active academic communities. EnergyPLAN provides integrated energy
systems simulation at the regional and national scale. TIMES,
MESSAGE-ix, and the broader IAM family provide national and global
scenario analysis capabilities with decades of institutional application
and established policy uptake. Each of these tools is internally
sophisticated. None of them, on its own, addresses the full analytical
chain from a site-level operational decision through a regional
infrastructure constraint to a national energy trajectory, under deep
uncertainty, in a form that is simultaneously governed, traceable, and
accountable to multiple stakeholders with different perspectives.

The framework proposed here is not a replacement for any of these tools.
It is an architecture within which they can contribute to that full
analytical chain. PyPSA belongs in the Regional Module, where its
network-aware dispatch optimisation produces the adequacy signals and
dual-variable price indicators that the site-level pathway evaluation
requires. OpenModelica belongs in the Facility Module, where its thermal
network simulation produces the high-fidelity operational detail that a
scheduling-grade dispatch model needs. TIMES-NZ belongs in the
FutureArtefact ensemble calibration layer, where its national scenario
outputs provide the macroeconomic and policy boundary conditions within
which regional and site-level futures are situated. The framework tells
each tool what it is responsible for producing, in what form, and with
what provenance. It does not tell each tool how to produce it. The
internal sophistication of each tool is liberated rather than
constrained by the architecture.

The practical value of the surrogate development pathway deserves
specific emphasis. A full PyPSA regional network model for the Southland
context is computationally expensive to run. Running it inside a DMDU
ensemble of several hundred or thousand futures is currently
impractical. Training a decision-ranked ML surrogate of that model,
following the validation protocol described in Appendix I-C, makes the
ensemble expansion tractable without sacrificing the decision-ranking
fidelity that the comparison requires. The surrogate does not replace
PyPSA; it emulates PyPSA's input-output behaviour at the level of the
decision-relevant outputs the framework requires, and the full PyPSA
model remains available for validation and for cases near regime
boundaries where the surrogate's confidence score falls below the
declared threshold. The same surrogate architecture applies to a
spatially explicit biomass logistics model, to a thermal process
simulation, and to a simplified coupling with macroeconomic scenario
generators such as TIMES-NZ. At every generation of implementation, what
crosses the module boundary is the same schema-conforming artefact. The
analytical chain does not need to know which generation produced it.

For modellers working with the IEA Project BlueSky interoperability
initiative, the framework offers a complementary governance layer.
BlueSky's ambition of standardised interfaces between heterogeneous
modelling components is architecturally aligned with the thin-waist
principle developed here. The additional contribution of the present
framework is artefact governance: the schema conformance, provenance
declaration, validation gating, and lineage tracking that make modelling
contributions not only interoperable but auditable across the full
analytical chain.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{It asks corporate decision-makers to recognise that the
analytical environment they need for long-horizon investment decisions
under deep uncertainty already exists in fragmentary form inside their
organisations, and that the framework's contribution is to connect those
fragments in a governed, coherent way that creates value they cannot
currently
access.}\label{it-asks-corporate-decision-makers-to-recognise-that-the-analytical-environment-they-need-for-long-horizon-investment-decisions-under-deep-uncertainty-already-exists-in-fragmentary-form-inside-their-organisations-and-that-the-frameworks-contribution-is-to-connect-those-fragments-in-a-governed-coherent-way-that-creates-value-they-cannot-currently-access.}

Large industrial companies managing long-lived asset portfolios in a
decarbonising economy are not operating without analytical support. They
have energy management systems that record real-time and historical
operational data at hourly or sub-hourly resolution. They have financial
models that project operating costs under different fuel price and
carbon price scenarios. They have engineering assessments that
characterise the technical performance of alternative technology
pathways. They have sustainability reporting systems that aggregate
emissions data across facilities. They have grid connection assessments
that characterise the available hosting capacity at their points of
connection. What they typically lack is an analytical architecture that
connects these data sources through governed interfaces, evaluates the
pathways they reveal under structured uncertainty across multiple
futures, and produces decision-relevant metrics that are simultaneously
precise enough to be actionable and honest enough to be defensible under
external scrutiny.

The framework is designed to receive the data that already exists within
industrial organisations and translate it into governed artefacts. An
existing energy management system's operational data becomes the basis
for a calibrated DemandPack rather than a synthetic one. An existing
engineering assessment of boiler replacement costs becomes the capital
cost parameter in the site module's configuration artefact. An existing
financial model's fuel price and carbon price projections become the
calibration basis for the FutureArtefact ensemble's uncertain driver
ranges. An existing grid connection assessment becomes the reference
capacity estimate around which the headroom multiplier is scaled. None
of this requires the organisation to start from scratch. It requires the
existing analytical work to be translated into the framework's artefact
format, which is the precondition for connecting it to the regional and
policy layers that the site-level analysis alone cannot address.

The commercial advantage this offers is specific and demonstrable. A
corporate investment decision made on the basis of a governed,
traceable, multi-layer robustness analysis is a qualitatively different
document from one made on the basis of a point-estimate techno-economic
assessment. It can demonstrate to the board that the pathway
recommendation remains defensible under a declared range of future
conditions rather than only under central-case assumptions. It can
demonstrate to investors that the capital allocation decision has been
stress-tested against the scenarios that ESG-focused investment
frameworks now require. It can demonstrate to regulators that the
company has evaluated the system-level consequences of its pathway
choice and has not simply optimised within its own boundary at the
expense of shared infrastructure resources.

For a company managing multiple sites, the multi-site extension offers a
further commercial advantage: the ability to evaluate interaction
effects between sites explicitly. The aggregate electricity demand of
multiple electrification pathway commitments across a South Island
portfolio is not the sum of individual site assessments. It is a
system-level demand profile that hits multiple GXPs simultaneously,
competes for reinforcement investment prioritisation by Transpower, and
creates aggregate biomass demand that may constrain the supply available
to any individual site. The corporate-scale deployment of the framework
makes these interactions visible and includes them in the pathway
comparison. A corporate sustainability officer who can demonstrate that
a portfolio transition strategy has been evaluated against system-level
interaction effects occupies a significantly stronger position in
negotiations with energy regulators, network operators, and
co-investment agencies than one who cannot.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{It asks policy-makers and regulators to recognise that the
system-level incentive misalignments revealed by the framework's
multi-perspective evaluation are not academic abstractions but
consequences of real, specific architectural features of the current
policy and pricing environment, features that can be identified
precisely, attributed to specific cost components and pricing
arrangements, and addressed through targeted policy
design.}\label{it-asks-policy-makers-and-regulators-to-recognise-that-the-system-level-incentive-misalignments-revealed-by-the-frameworks-multi-perspective-evaluation-are-not-academic-abstractions-but-consequences-of-real-specific-architectural-features-of-the-current-policy-and-pricing-environment-features-that-can-be-identified-precisely-attributed-to-specific-cost-components-and-pricing-arrangements-and-addressed-through-targeted-policy-design.}

The framework's most important policy finding to date is the divergence
between site-perspective and system-perspective cost assessments for the
electrification pathway under GXP-constrained futures. Under current
network pricing arrangements, the cost of grid reinforcement triggered
by a new large industrial load may not be directly charged to the load
that triggered it, creating a systematic incentive for individual site
operators to pursue electrification pathways whose private cost appears
competitive even when the full-system cost is significantly higher. This
misalignment is not visible in any single-site techno-economic
assessment. It is only visible when the site's incremental electricity
demand is evaluated against the regional grid system's actual hosting
capacity constraints, and when the cost of the resulting reinforcement
requirement is attributed from both the private and public perspectives
simultaneously. The framework makes this attribution explicit, and it
attributes it to specific futures and specific cost components, so that
the policy question becomes precise: under which conditions do the
misalignment occur, how large is it, and which policy instrument most
directly addresses it?

For EECA, the framework offers a complement to the RETA screening
methodology that addresses exactly the limitation that RETA's current
approach cannot reach: the time-resolved, uncertainty-aware,
multi-perspective evaluation of individual site pathway decisions
against regional infrastructure constraints. RETA identifies where the
decarbonisation opportunities are and what infrastructure would be
needed to realise them. The framework evaluates whether those
opportunities remain robust across the futures in which the
infrastructure may or may not materialise on the required timeline. The
two approaches are designed for each other.

For the Climate Change Commission, the framework offers a tool for
evaluating whether the sectoral emissions reduction targets in the
emissions reduction plan are achievable under the specific
infrastructure and supply chain conditions that determine feasibility at
the site level. A national emissions reduction plan that requires
large-scale industrial electrification without representing the regional
grid infrastructure constraints that determine feasibility on the
required timeline may look achievable in a national model and be
infeasible on the ground. The framework's site-to-region coupling makes
that gap analytically visible and attributable.

For the Electricity Authority and Transpower, the framework offers a
structured approach to the aggregate demand forecasting challenge that
large-scale industrial electrification creates. Multiple industrial
users electrifying simultaneously in the same region creates aggregate
GXP demand that is difficult to predict from individual site assessments
alone. A framework-based regional demand assessment, aggregating
IncrementalElectricityPacks from multiple facility modules through a
shared regional module, produces an aggregate demand forecast that is
grounded in individual facility operational characteristics, evaluated
across a structured future ensemble, and expressed in the time-resolved
format that network investment planning requires.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{It asks communities and civil society organisations to
recognise that the decisions being made by large industrial operators
about the energy infrastructure of the next thirty years are
simultaneously decisions about the regional economy, the regional
employment base, the regional environment, and the social fabric of
communities built around particular industrial
configurations.}\label{it-asks-communities-and-civil-society-organisations-to-recognise-that-the-decisions-being-made-by-large-industrial-operators-about-the-energy-infrastructure-of-the-next-thirty-years-are-simultaneously-decisions-about-the-regional-economy-the-regional-employment-base-the-regional-environment-and-the-social-fabric-of-communities-built-around-particular-industrial-configurations.}

These consequences are not secondary effects that happen after the real
decision has been made. They are part of the decision. An analytical
environment that treats them as qualitative footnotes to a financial
analysis produces a decision that is less likely to be legitimate, less
likely to be durable, and less likely to serve the communities that must
live with it.

The framework's multi-perspective evaluation architecture is designed to
make social and environmental consequences first-class objects in the
comparison chain alongside cost and emissions. A regional employment
impact module can be added to the regional layer, producing an
employment consequence artefact alongside the SignalsPack. A community
infrastructure stress module can evaluate whether the regional
development patterns associated with different industrial transition
pathways are consistent with community planning priorities. A
biodiversity and land-use consequence module can represent the
environmental implications of biomass supply chain development at the
scale where those implications are most material.

Each of these modules follows the same thin-waist architecture as the
electricity and cost modules. It receives a standardised input artefact,
produces a standardised output artefact, and contributes to the
DecisionSummaryArtefact a consequence dimension that can be evaluated
through the same regret and robustness metrics as the financial
dimensions. This is not social accounting tacked onto an economic model.
It is the decision-first principle applied to the full scope of what the
decision is actually about. When a regional community has a stake in
whether a large employer pursues electrification or biomass conversion,
that stake is a decision-relevant consequence. It belongs in the
analysis from the beginning.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{It asks AI assistants and the research community to engage
with this document not as a completed account of a finished framework
but as the first formal specification of a knowledge environment that is
designed to grow through exactly the kind of engagement that
AI-assisted, community-driven research can
provide.}\label{it-asks-ai-assistants-and-the-research-community-to-engage-with-this-document-not-as-a-completed-account-of-a-finished-framework-but-as-the-first-formal-specification-of-a-knowledge-environment-that-is-designed-to-grow-through-exactly-the-kind-of-engagement-that-ai-assisted-community-driven-research-can-provide.}

The framework's modular architecture is the blueprint for its own
extension. A domain specialist in water infrastructure planning who
reads this document and recognises the decision structure they are
working with can contribute a domain instantiation by specifying the
relevant module classes, artefact families, and uncertain drivers in
terms of the framework's interface contract. A PyPSA developer who reads
the regional module specification in Appendix III-F can implement that
module and contribute it to the community through the process described
in the Context Declaration. A researcher in the DMDU community who
identifies a new robustness metric or a new scenario discovery algorithm
that improves the evaluation layer can propose its integration as a
methodological extension. Each of these contributions enters a governed
environment: it is assessed for compatibility with the thin-waist
architecture and the decision-first principle, reviewed through the
editorial process described in the Context Declaration, and if accepted,
formally incorporated with full provenance.

The AI-assisted engagement pathway deserves specific mention. A
practitioner who uploads this document to a capable AI assistant and
asks questions drawn from the audience-specific catalogues is using the
document in exactly the way it was designed to be used: as a structured
knowledge base that supports precise, framework-grounded engagement with
the specific question at hand. The AI assistant is not being asked to
summarise the document. It is being asked to reason about the framework
with the precision that the framework's own vocabulary supports, to
distinguish between what has been implemented and what has been
specified, to help the user navigate from their specific question to the
part of the framework most relevant to addressing it, and to acknowledge
explicitly when a question concerns something the current framework does
not yet represent.

The framework does not need to be complete to be useful. It needs to be
honest about what it does and does not yet represent, to provide a clear
path toward representing what it does not, and to create the conditions
under which the research community and the communities of practice most
directly affected by the decisions the framework supports can contribute
to that path in a governed, accountable, and cumulatively productive
way. This document is the beginning of that path.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{A Note on How the AI Assistant Should Use This
Document}\label{a-note-on-how-the-ai-assistant-should-use-this-document}

When answering any question using this document, the AI assistant should
ground answers in the framework's specific terms and concepts rather
than in general knowledge about energy systems or decision theory that
may differ from how this framework uses those terms. It should
acknowledge explicitly when a question asks about something the current
framework does not yet implement and explain what would be required to
address it, consistent with the progressive-refinement philosophy. It
should distinguish between what the Edendale proof of concept has
demonstrated, what has been specified as next-phase development, and
what remains a longer-range vision.

The assistant should treat the module interface declarations at the
front of each Part as the authoritative map of what each Part
contributes and what it assumes. It should use the numbered propositions
in Part IV Chapter 9 as the distilled claim set when asked about the
framework's intellectual contribution. It should treat the
decision-first boundary principle as the conceptual anchor that connects
every architectural choice to the framework's foundational commitment.
When a question concerns a domain outside process heat, the assistant
should use the framework's architecture as the basis for reasoning about
how that domain would be instantiated, specifying the relevant decision
class, the relevant module structure, and the relevant artefact families
in terms the framework's interface contract supports.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{What This Document Contains}\label{what-this-document-contains}

The document is organised across four Parts and nine chapters, preceded
by a preliminary orientation Part (Part 0). Every Part opens with a
module interface declaration specifying what the Part assumes, what it
contributes, what it passes forward, and whether it can be read
independently.

Part 0 provides the framework map, the decision-first boundary
principle, the seven functional layers, the current implementation
status, and the navigational logic for the rest of the document.

Part I establishes the conceptual and methodological foundations: why
decision-first boundary setting is the appropriate response to
long-horizon planning under deep uncertainty, what analytical grammar is
needed to pursue it rigorously, and what tools are available to generate
decision-relevant consequences.

Part II develops the architecture: modular decomposition, thin-waist
artefact exchange, and the governed data backbone that makes the
environment persistent, traceable, and extensible.

Part III instantiates the architecture: first in the domain of coupled
energy systems and industrial process heat, then in the specific New
Zealand conditions that motivate the proof of concept, and finally in
the full documented Edendale--Southland proof of concept with its
implemented pipeline, its artefact integrity record, and its results.

Part IV synthesises extending the framework beyond its current
instantiation, stating its limits honestly, and presenting the eleven
propositions that constitute its intellectual claim.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Current Implementation
Status}\label{current-implementation-status}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Component
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Status
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Significance
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
DemandPack construction (synthetic, RETA-calibrated) & Implemented &
Foundational demand artefact for all epochs \\
Proportional and optimal-subset dispatch & Implemented & Site-level
pathway evaluation with cost and emissions accounting \\
SignalsPack generator (Edendale GXP, schema v0.1.0) & Implemented &
Frozen interface contract with SHA256 hash integrity \\
GXP screening module (stylised) & Implemented & First-generation
regional adequacy assessment \\
Grid RDM evaluation (21 to 100 futures) & Implemented & Robustness
screening across five uncertain driver dimensions \\
Site decision robustness overlay & Implemented & Site-cost uncertainty
separated from grid-side uncertainty \\
Pathway comparison (2035\_EB vs 2035\_BB) & Implemented & Paired-futures
regret and robustness diagnostics \\
DuckDB and Parquet analytical backbone & Specified & Governed,
append-only, queryable persistence layer \\
PyPSA regional electricity module & Specified & Full network
optimisation replacing stylised screen \\
OpenModelica site truth model & Specified & Equation-based physical
simulation replacing proportional dispatch \\
Surrogate-accelerated ensemble & Specified & ML emulator enabling
ensemble expansion to thousands of futures \\
TIMES-NZ scenario coupling & Vision & National-regional consistency in
FutureArtefact ensemble calibration \\
Corporate-scale multi-site deployment & Vision & Portfolio transition
strategy across multiple facilities \\
Multi-domain extension & Vision & Water, transport, and land-use
instantiations through community contribution \\
Natural-language query interface & Vision & Accountability-grade
interrogation of the governed artefact store \\
\end{longtable}
}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Key Concepts Defined Here}\label{key-concepts-defined-here}

\textbf{Decision-first boundary principle.} The foundational commitment
of this framework: that analytical boundaries are drawn primarily by the
decision context and the consequences that must remain visible for
comparison, not by the physical extent of the system being studied.
Physical parameters, engineering constraints, and process detail matter
and are incorporated, but they do not determine the architecture. What
determines the architecture is the decision: what is being chosen, by
whom, with consequences at which scales, and under which dimensions of
uncertainty. The framework develops toward the parameters that influence
the decision, not toward the completeness of the system description. The
decision makes the boundary; the artefact makes the connection.

\textbf{Decision-centred modelling.} An approach in which the analytical
environment is organised around the requirements of decision
exploration, specifically alternatives, futures, consequences, and
metrics, rather than around the internal logic of any one model class.
The decision problem is the primary object of analysis; models are
instruments in service of that analysis.

\textbf{Progressive decision focus.} The development philosophy that
follows from the decision-first boundary principle. At each stage of
framework development, the question is: what is the most consequential
simplification currently in the model, in the sense that removing it
would most change the decision-relevant comparison? The answer, produced
by regret sensitivity diagnostics and scenario discovery, determines
what is developed next. Complexity is added where it improves decision
quality, not where it increases formal sophistication.

\textbf{Multi-layer decision environment.} An analytical environment in
which multiple spatial, temporal, and institutional scales are
simultaneously represented, with consequences flowing between scales
through governed artefact exchanges. The multi-layer structure follows
from the decision-first principle: many decisions that appear to be
single-layer have consequences at multiple layers, and representing only
one layer produces a systematically incomplete analysis.

\textbf{Deep uncertainty.} Conditions in which stakeholders cannot agree
on the appropriate models, the probability distributions of key
parameters, or the metrics by which outcomes should be judged. Deep
uncertainty is not a deficiency of data. It is a structural feature of
certain planning environments and requires different analytical tools
from risk, where probabilities can be estimated credibly.

\textbf{Module.} A component of the analytical environment with a
defined role, declared inputs, declared outputs, and explicit interface
contracts. A module may be implemented in different ways without
changing its analytical role, provided it continues to emit the required
artefacts.

\textbf{Thin waist.} The principle that what is exchanged between
modules must be narrow, explicit, and stable, while what happens inside
each module may evolve freely. The thin waist separates the stability of
the interface from the flexibility of the implementation. It is a design
principle, not a data format.

\textbf{Artefact.} A governed analytical object: a structured,
versioned, validated, and provenance-carrying output that enters the
comparison chain. An artefact is not merely a file or a model output. It
is an output that has been materialised in a schema-conforming form,
linked to the run and future conditions that produced it, passed through
a validation gate, and stored in the analytical backbone with explicit
lineage.

\textbf{Artefact families.} The canonical types of artefact the
framework uses. The primary families are: DemandPack (time-resolved site
heat demand); IncrementalElectricityPack (GXP-facing incremental
electricity demand signal); SignalsPack (GXP-level electricity interface
signals from the regional module); FutureArtefact (the uncertain-driver
combination under which a run was executed); ResultArtefact (site-level
cost, emissions, and adequacy outcomes); ValidationArtefact (acceptance
gate records); and DecisionSummaryArtefact (regret, robustness, and
threshold metrics across the ensemble).

\textbf{DemandPack.} The canonical demand artefact. A time-resolved,
versioned, acceptance-gated record of site-level heat demand for a
specific epoch, scenario variant, and future condition.

\textbf{IncrementalElectricityPack.} The canonical electricity interface
artefact produced by the site dispatch module and consumed by the
regional electricity module. It carries the compact signal descriptors
representing the site's electricity demand increment at the GXP
boundary: annual incremental energy, peak incremental demand, 95th
percentile demand, winter-peak share, and shape cluster label. It is the
thin-waist object at the site-to-regional boundary.

\textbf{SignalsPack.} The canonical electricity interface artefact
produced by the regional module and consumed by the site module. A
structured set of GXP-level indicators including adequacy headroom,
tariff adders, scarcity-price proxies, upgrade-class signals, and
confidence scores. In the current proof of concept, produced by the
Edendale GXP repository using frozen schema version 0.1.0 with SHA256
hash integrity.

\textbf{FutureArtefact.} The canonical uncertainty artefact. Records the
specific combination of uncertain driver values under which a given run
was executed, including the headroom multiplier, the regional demand
growth multiplier, the hydro year class, the biomass availability
multiplier, the biomass cost multiplier, and the ETS carbon price.

\textbf{DMDU (Decision Making under Deep Uncertainty).} The intellectual
tradition that reorients analytical practice from optimising against a
single forecast toward characterising how decisions perform across a
wide ensemble of plausible futures. Key methods include Robust Decision
Making (RDM), Dynamic Adaptive Policy Pathways (DAPP), Exploratory
Modelling and Analysis (EMA), Many-Objective Robust Decision Making
(MORDM), and scenario discovery methods including PRIM and CART.

\textbf{RDM (Robust Decision Making).} A DMDU method developed
principally at RAND by Lempert, Popper, and Bankes that evaluates
alternatives across large ensembles of futures using regret, satisficing
frequency, and vulnerability diagnostics rather than expected-value
optimisation.

\textbf{Regret.} The difference between what a chosen strategy achieves
under a given future and what the best available strategy would have
achieved under that same future. Regret is future-specific and
comparative.

\textbf{Robustness.} A property of a strategy that performs acceptably
or competitively across many plausible futures. A robust strategy may
not be optimal under any single future; its value lies in avoiding large
failures across many futures.

\textbf{Satisficing.} A decision standard, following Herbert Simon, that
asks which alternatives meet stated thresholds of acceptability across
futures rather than which is nominally best.

\textbf{GXP (Grid Exit Point).} In the New Zealand electricity system,
the point at which electricity leaves the national transmission grid.
GXP-level hosting capacity is the primary constraint governing
electrification feasibility in the Southland regional context.

\textbf{The Edendale--Southland PoC (Proof of Concept).} The framework's
primary instantiation. Edendale is Fonterra's dairy processing facility
in Southland, New Zealand. The PoC implements a one-pass coupling chain
evaluating two 2035 pathways, 2035\_EB (Electrified Boiler) and 2035\_BB
(Biomass Boiler), across a paired ensemble of 21 to 100 futures. It is a
methodological demonstration, not a planning-grade forecast.

\textbf{One-pass coupling.} The coupling pattern implemented in the
current proof of concept: modules run in a defined sequence without
iterative feedback. The site dispatch module produces the incremental
electricity signal; the regional screening module evaluates grid
constraints; results are aggregated without looping.

\textbf{Site decision robustness overlay.} A post-processing layer that
adds stylised site-cost uncertainty to the grid RDM evaluation, enabling
non-trivial robustness analysis of the pathway decision without breaking
the one-pass coupling architecture. Applies triangular distributions for
electricity price, biomass price, and ETS multipliers across the paired
futures.

\textbf{The analytical backbone.} The governed persistence layer in
which all artefacts are stored, linked, and made queryable. The backbone
is append-only, schema-governed, and lineage-linked. In the current
proof of concept, implemented as a file-system run-bundle structure with
SHA256 hash integrity; the DuckDB and Parquet implementation is the
next-phase target.

\textbf{Progressive refinement.} The framework's development philosophy.
A minimal implementation is architecturally complete and analytically
useful before any of the richer components are built. The architecture
reveals its own next steps through regret sensitivity diagnostics.
Complexity is added where it improves decision quality.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Community Extension}\label{community-extension}

The document is released as a living resource. The framework's modular
architecture is designed to accommodate extension without disruption: a
new domain instantiation adds a new Part-equivalent appendix; a new
artefact family adds a new schema entry; a new uncertain driver adds a
field to the FutureArtefact; a new regional application adds a new
proof-of-concept chapter.

Contributions are welcomed in four categories. Domain instantiations
apply the framework to domains beyond industrial process heat. Each
instantiation should specify the relevant decision class, module
structure, and artefact families in terms of the framework's interface
contract. Module implementations provide alternative implementations of
any module satisfying the same interface contract. Methodological
extensions extend the evaluation layer with new robustness metrics,
scenario discovery approaches, or adaptive pathway formulations.
Corrections and critiques identify errors or propose alternative
approaches.

Contributions may be submitted through the GitHub repository associated
with this document, using the structured templates provided in the
GitHub Discussions space. The editorial process assesses compatibility
with the thin-waist architecture and the decision-first principle before
formal incorporation.

\chapter{Acknowledgement of AI-Assisted
Development}\label{acknowledgement-of-ai-assisted-development}

This framework and this document were developed in active and sustained
collaboration with AI tools throughout. Research synthesis, code
scaffolding, schema drafting, iterative prototyping, artefact validation
logic, and significant portions of the writing were produced through
dialogue with AI assistants, principally Claude (Anthropic), ChatGPT
(OpenAI), Grok, and Cursor AI, across the full duration of the research
process.

This is acknowledged here not as a disclaimer but as a methodological
fact with direct relevance to the framework's own argument. The
framework argues that any output entering an analytical comparison chain
must be schema-conforming, provenance-carrying, and validation-gated,
regardless of whether it was produced by a deterministic optimiser, a
stochastic simulation, or an AI-assisted development process. This
document's own production is subject to the same governance logic it
prescribes. Without governance, AI-assisted analytical work is opaque.
With it, it is traceable. This document is designed to make that
boundary visible.

Readers who are themselves using AI tools to engage with this framework
are doing something it explicitly anticipates and accommodates. This
document is designed to be an evolving, AI-readable, progressively
extensible knowledge environment. This is its current state. Future
states will be richer. The architecture is designed to absorb that
enrichment without losing coherence.

\chapter{Preamble}\label{preamble}

How to Read This Document

\hfill\break

There are several ways to come to this document, and it has been written
to be useful from all of them. But before describing the reading paths,
it is worth saying something about what the document is trying to do at
the broadest level, because the ambition is larger than the proof of
concept might suggest.

\section{The Ambition}\label{the-ambition}

Complex planning decisions in the early twenty-first century are made
under conditions that conventional analytical tools were not designed to
handle. The decisions are long-horizon, their consequences propagate
across scales and actors, the futures that determine their quality
cannot be agreed upon probabilistically, and the stakeholders who must
live with the outcomes span private corporations, public regulators,
regional communities, and global environmental accounting systems. None
of the tools currently available to support such decisions was designed
to address all of these characteristics simultaneously. National energy
system models address long horizons and cross-sector consistency but
abstract away the operational heterogeneity of individual facilities and
the local infrastructure constraints that determine feasibility.
Site-level techno-economic assessments address operational realism but
cannot see beyond the facility boundary. Regulatory impact assessments
address stakeholder perspectives but rarely achieve the analytical
precision needed to evaluate pathway robustness under deep uncertainty.
The result is that consequential decisions about the energy
infrastructure of the next thirty years are routinely made on the basis
of analyses that are each individually sophisticated but collectively
incoherent, because they are not connected through a common architecture
that allows their contributions to be evaluated together.

The framework developed in this document is a response to that
incoherence. It is not a better model. It is an architecture within
which better models can contribute to a coherent analytical chain,
evaluated against a shared set of decision-relevant metrics, under a
shared representation of the futures across which the decision must
remain defensible.

\section{What It Could Become}\label{what-it-could-become}

The proof of concept demonstrated in Module 6 evaluates two pathways for
one facility against 64 futures under one analytical architecture. That
is a very small instantiation of what the framework's architecture could
support. The same architecture, deployed at the scale of a large
industrial company's full facility portfolio, coupled to a national
energy system model, equipped with a trained ML surrogate of the
regional electricity module allowing the ensemble to span several
thousand futures, and connected to a natural-language query interface
over the governed artefact store, would allow a regional planner, a
corporate sustainability officer, a policy analyst, or a community
representative to ask the analysis a specific question and receive an
answer grounded in specific artefacts whose provenance and validation
status are visible. That is not science fiction. Every component of that
vision is technically feasible with existing tools and methods. What
does not yet exist is the architecture that connects them in a governed,
traceable, decision-centred way. That architecture is what this document
describes.

\section{What It Asks}\label{what-it-asks}

The framework asks readers to accept a specific intellectual commitment
that runs against the grain of most modelling practice: that the
appropriate starting point for complex planning analysis is not the
system but the decision; that the model should grow toward the
parameters that influence the decision, not toward the completeness of
the physical description; and that the fog of deep uncertainty is not
cleared by more data or more computation alone but by a more disciplined
organisation of what is known, what is uncertain, and what must remain
visible for the comparison at hand. At each step of development, the
framework lights up the domain that needs to be explored next. It does
not promise to eliminate uncertainty. It promises to make decisions more
intelligible in its presence.

\section{On the Document's Own
Incompleteness}\label{on-the-documents-own-incompleteness}

This document is honest about what has been built and what has not. The
proof of concept in Module 6 is bounded in ways that are explicitly
declared. Several modules are described as specifications rather than
implemented systems. Several sub-modules are roadmaps for future
development rather than accounts of existing work. This incompleteness
is not a deficiency. It is a feature of the progressive-refinement
philosophy: a bounded, auditable starting point that reveals its own
next steps is more valuable than a complete but opaque system. The
architecture is designed to absorb enrichment. This document is designed
to be extended.

\section{The Reading Paths}\label{the-reading-paths}

If you are new to the framework, begin with Module 0 (Orientation),
which provides the full module map and the three-level architecture in
overview form. If you are a practitioner with a specific question, use
the Node Declaration Tables at the front of each section and the
question catalogues in the Context Declaration to navigate directly. If
you want the conclusions before the argument, read Module 7 (Extension
and Synthesis) first. If you want to know what has been built and how to
run it, read Module 6 and its sub-modules. If you are an AI assistant
that has been given this document, the Context Declaration is addressed
to you directly.

\chapter{Navigation Guide}\label{navigation-guide}

The Node Declaration System Explained

\hfill\break

This document is structured so that both human readers and AI agents can
navigate it efficiently, processing what is relevant to their purpose
and skipping what is not. The mechanism that makes this possible is the
\textbf{Node Declaration Table}, which appears at the start of every
section and sub-module in the manuscript. This guide explains how to
read it and how to use it.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{1. The Node Declaration
Table}\label{the-node-declaration-table}

Every section and sub-module opens with a table in the following format:

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, bottomrule=.15mm, bottomtitle=1mm, breakable, colback=white, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, coltitle=black, left=2mm, leftrule=.75mm, opacityback=0, opacitybacktitle=0.6, rightrule=.15mm, title={Node Declaration --- §X.X Section Title}, titlerule=0mm, toprule=.15mm, toptitle=1mm]

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Field
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Content
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Tier} & Section or Sub-Module \\
\textbf{Status} & Implemented / Specified / Vision \\
\textbf{Assumes} & Prior sections the reader must know \\
\textbf{Contributes} & One sentence: what this node adds to the
argument \\
\textbf{Skip condition} & Precise statement of when this node can be
skipped \\
\textbf{Passes to} & Subsequent sections that depend on this one \\
\textbf{Sub-Modules here} & Sub-modules that follow this section, or
None \\
\end{longtable}
}

\end{tcolorbox}

Each field has a specific function.

\textbf{Tier} distinguishes a main-body Section from a Sub-Module.
Sections carry the primary argument. Sub-Modules carry technical depth
that supports the argument but is not required to follow it.

\textbf{Status} declares the current state of the content this node
covers. \emph{Implemented} means the described capability exists and is
operational in the proof of concept. \emph{Specified} means the
capability has been fully designed but not yet built. \emph{Vision}
means the capability has been identified as a future direction but is
not yet specified in detail.

\textbf{Assumes} lists the prior sections a reader must have processed
for this node to make sense. If the listed sections are already known
from prior reading or professional background, the reader may enter here
without reading them first.

\textbf{Contributes} is a single sentence stating what this node adds.
It answers the question: what will I understand after reading this that
I did not understand before? It is the primary navigation aid for both
human readers and AI agents assessing whether to engage with the full
node.

\textbf{Skip condition} is a precise statement of when this node may be
passed over. It is written to be testable: a reader can assess whether
the condition applies to them and act accordingly. Skip conditions are
not invitations to skip carelessly; they are declarations of what prior
knowledge makes this node redundant.

\textbf{Passes to} lists the sections that directly depend on this node.
If you skip a section, check this field in the sections that follow it:
if your skipped section appears in their Assumes field, you may need to
return.

\textbf{Sub-Modules here} lists any sub-modules that immediately follow
this section. Sub-modules are always optional unless a downstream
section explicitly lists them in its Assumes field.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{2. The Skip Condition: Instructions for AI
Agents}\label{the-skip-condition-instructions-for-ai-agents}

An AI agent processing this document should apply the following rules.

If the skip condition is met, the agent may limit its processing to the
first paragraph of the section and the Node Declaration Table. The first
paragraph always states the section's core claim or function; the Node
Declaration Table provides the dependency structure. Together, these are
sufficient for the agent to maintain an accurate structural map of the
document without processing every section in full.

If the skip condition is not met, the agent must process the full
section before proceeding.

Sub-modules are always optional. No downstream section depends on a
sub-module unless that sub-module is explicitly listed in the downstream
section's Assumes field. An agent may skip all sub-modules in a first
pass and return to specific ones when a query requires the technical
depth they contain.

When answering a user query, the agent should first identify which
sections are relevant by consulting the Contributes and Passes to
fields, then process those sections fully. Sections outside this set may
be processed at summary level only.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{3. The Four Reading Paths}\label{the-four-reading-paths}

\textbf{Sequential argument.} Read every section in module order from
Module 0 through Module 7, consulting sub-modules where the text
references them. This path gives the fullest picture and is recommended
for researchers engaging with the methodological contribution.

\textbf{Modular reference.} Identify the section most relevant to your
question using the Table of Contents purpose annotations. Check that
section's Assumes field and read any listed prior sections you do not
already know. This path is recommended for practitioners, implementers,
and domain specialists.

\textbf{Framework overview.} Read Module 0 (Orientation) and Module 7
(Extension and Synthesis) only. Module 0 provides the full architectural
map; Module 7 provides the eleven propositions that constitute the
framework's intellectual claim. Together they give the argument's
beginning and end without the internal machinery.

\textbf{AI-assisted engagement.} Upload the full document to an AI
assistant. Use the audience-specific question catalogues in the Context
Declaration to begin. The AI assistant will use the Node Declaration
Tables to navigate the document precisely in response to your questions.
You may also ask the assistant to identify which sections are relevant
to a specific question before reading them yourself.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{4. The Sub-Module Structure}\label{the-sub-module-structure}

Sub-modules follow immediately after the section they deepen. They are
separated from the main section by a horizontal rule and open with their
own Node Declaration Table. Each sub-module has its own skip condition
and its own Tier field set to Sub-Module.

The purpose of this arrangement is to keep the main argument
uninterrupted. A reader following the primary argument encounters the
sub-module's Node Declaration Table, assesses its skip condition, and
either proceeds into the sub-module or returns to the main sequence. The
argument does not require the sub-module; the sub-module does not
require the reader to have read the section in any particular way before
entering it.

Sub-module identifiers follow the pattern SM-X.X-Y where X.X is the
parent section number and Y is an alphabetical letter distinguishing
multiple sub-modules within the same section.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{5. The Table of Contents}\label{the-table-of-contents}

Every entry in the Table of Contents carries a one-line purpose
annotation in italics. This annotation is the Contributes field stated
even more briefly. Status indicators appear after each entry: ✓ for
Implemented, ○ for Specified, and ◇ for Vision. Sub-modules are indented
one level below their parent section and marked with the prefix symbol ⌐
to distinguish them visually from main-body sections.

Example Table of Contents entries: Module 3: Architecture
\ldots\ldots\ldots\ldots\ldots\ldots\ldots\ldots\ldots\ldots\ldots\ldots\ldots\ldots\ldots\ldots..
45 §3.5 Artefact Families and Schema Design ✓ Seven canonical families
and five schema design principles \ldots{} 52 ⌐ SM-3.5-A Canonical
Artefact Schema Reference ✓ Complete field-level specifications for all
families \ldots\ldots.. 58 §3.6 Artefact Lifecycle, Validation, and
Acceptance Gates ✓ From provisional generation through validated
publication \ldots.. 65

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{6. Worked Example}\label{worked-example}

\textbf{Hypothetical query:} ``I need to understand what fields the
SignalsPack artefact carries and how its schema is validated.''

The agent consults the Table of Contents and identifies two candidates:
§3.5 (Artefact Families and Schema Design) and SM-3.5-A (Canonical
Artefact Schema Reference).

The agent reads the Node Declaration for §3.5:

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, bottomrule=.15mm, bottomtitle=1mm, breakable, colback=white, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, coltitle=black, left=2mm, leftrule=.75mm, opacityback=0, opacitybacktitle=0.6, rightrule=.15mm, title={Node Declaration --- §3.5 Artefact Families and Schema Design}, titlerule=0mm, toprule=.15mm, toptitle=1mm]

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Field
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Content
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Tier} & Section \\
\textbf{Status} & ✓ Implemented \\
\textbf{Assumes} & §3.4 (The Thin-Waist Artefact Exchange) \\
\textbf{Contributes} & Defines the seven canonical artefact families and
establishes the five schema design principles governing their
structure \\
\textbf{Skip condition} & Skip if reader already knows the seven
artefact family names, their primary producers and consumers, and the
meaning of schema version stability \\
\textbf{Passes to} & §3.6, SM-3.5-A, Module 6 \\
\textbf{Sub-Modules here} & SM-3.5-A: Canonical Artefact Schema
Reference \\
\end{longtable}
}

\end{tcolorbox}

The skip condition is not met because the query requires specific
field-level knowledge (SignalsPack schema and validation), which the
skip condition does not cover. The agent processes §3.5 in full.

The agent then reads the Node Declaration for SM-3.5-A:

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, bottomrule=.15mm, bottomtitle=1mm, breakable, colback=white, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, coltitle=black, left=2mm, leftrule=.75mm, opacityback=0, opacitybacktitle=0.6, rightrule=.15mm, title={Node Declaration --- SM-3.5-A: Canonical Artefact Schema Reference}, titlerule=0mm, toprule=.15mm, toptitle=1mm]

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Field
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Content
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Tier} & Sub-Module \\
\textbf{Status} & ✓ Implemented \\
\textbf{Assumes} & §3.5 \\
\textbf{Contributes} & Complete field-level specifications for all seven
artefact families including column names, data types, units, and
validation rules \\
\textbf{Skip condition} & Skip if only the conceptual definition of
artefact families is needed; process in full if implementing the
framework or verifying schema compliance \\
\textbf{Passes to} & Module 6 §6.4 (implementation uses these schemas
directly) \\
\textbf{Sub-Modules here} & None \\
\end{longtable}
}

\end{tcolorbox}

The query requires the SignalsPack field-level specification and its
validation rules. The skip condition for SM-3.5-A is not met. The agent
processes SM-3.5-A in full and retrieves the SignalsPack schema fields
and their validation rules from the table within it.

The agent does not need to process §3.6, Module 4, or Module 5 to answer
this query.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, bottomrule=.15mm, bottomtitle=1mm, breakable, colback=white, colbacktitle=quarto-callout-important-color!10!white, colframe=quarto-callout-important-color-frame, coltitle=black, left=2mm, leftrule=.75mm, opacityback=0, opacitybacktitle=0.6, rightrule=.15mm, title=\textcolor{quarto-callout-important-color}{\faExclamation}\hspace{0.5em}{Important}, titlerule=0mm, toprule=.15mm, toptitle=1mm]

This Navigation Guide is itself governed by a Node Declaration. It has
no skip condition: it should be read by all users before engaging with
the manuscript's main body.

\end{tcolorbox}

\part{Module 0: Orientation}

\chapter{Module 0: Orientation}\label{module-0-orientation-1}

The Framework at a Glance

\hfill\break

\part{Module 1: Foundations}

\chapter{Module 1: Foundations}\label{module-1-foundations-1}

The Decision Problem and Its Requirements

\hfill\break

\chapter{Sub-Module 1.1-A}\label{sub-module-1.1-a}

Philosophical Lineage of Deep Uncertainty

\hfill\break

\chapter{Sub-Module 1.4-A}\label{sub-module-1.4-a}

Decision Theory

\hfill\break

\chapter{Sub-Module 1.7-A}\label{sub-module-1.7-a}

Requirements Mapping

\hfill\break

\part{Module 2: Modelling Foundations}

\chapter{Module 2: Modelling
Foundations}\label{module-2-modelling-foundations-1}

Tools, Machinery, and the Logic of Exploration

\hfill\break

\chapter{Sub-Module 2.2-A}\label{sub-module-2.2-a}

Optimisation Roles

\hfill\break

\chapter{Sub-Module 2.3-A}\label{sub-module-2.3-a}

Surrogate Validation

\hfill\break

\chapter{Sub-Module 2.4-A}\label{sub-module-2.4-a}

Scenario Discovery

\hfill\break

\part{Module 3: Architecture}

\chapter{Module 3: Architecture}\label{module-3-architecture-1}

Modular Decomposition, Artefacts, and Governed Persistence

\hfill\break

\chapter{Sub-Module 3.5-A}\label{sub-module-3.5-a}

Artefact Schema

\hfill\break

\chapter{Sub-Module 3.7-A}\label{sub-module-3.7-a}

Backbone Implementation

\hfill\break

\chapter{Sub-Module 3.8-A}\label{sub-module-3.8-a}

AI/ML Governance

\hfill\break

\part{Module 4: Domain Translation}

\chapter{Module 4: Domain
Translation}\label{module-4-domain-translation-1}

Energy Systems, Process Heat, and Coupled Evaluation

\hfill\break

\chapter{Sub-Module 4.4-A}\label{sub-module-4.4-a}

System Regret

\hfill\break

\part{Module 5: The New Zealand Context}

\chapter{Module 5: The New Zealand
Context}\label{module-5-the-new-zealand-context-1}

Conditions, Constraints, and Empirical Grounding

\hfill\break

\chapter{Sub-Module 5.2-A}\label{sub-module-5.2-a}

GXP Capacity

\hfill\break

\chapter{Sub-Module 5.3-A}\label{sub-module-5.3-a}

Biomass Supply

\hfill\break

\chapter{Sub-Module 5.4-A}\label{sub-module-5.4-a}

ETS History

\hfill\break

\part{Module 6: The Proof of Concept}

\chapter{Module 6: The Proof of
Concept}\label{module-6-the-proof-of-concept-1}

Architecture, Implementation, and Results

\hfill\break

\chapter{Sub-Module 6.2-A}\label{sub-module-6.2-a}

PoC Pipeline

\hfill\break

\chapter{Sub-Module 6.3-B}\label{sub-module-6.3-b}

DemandPack Methods

\hfill\break

\chapter{Sub-Module 6.4-C}\label{sub-module-6.4-c}

Site Interface Contract

\hfill\break

\chapter{Sub-Module 6.5-D}\label{sub-module-6.5-d}

SignalsPack Contract

\hfill\break

\chapter{Sub-Module 6.6-E}\label{sub-module-6.6-e}

PyPSA Module Design

\hfill\break

\chapter{Sub-Module 6.8-F}\label{sub-module-6.8-f}

Validation Records

\hfill\break

\part{Module 7: Extension and Synthesis}

\chapter{Module 7: Extension and
Synthesis}\label{module-7-extension-and-synthesis-1}

Beyond Process Heat and the Closing Argument

\hfill\break

\bookmarksetup{startatroot}

\chapter{Glossary}\label{glossary}

Artefact Families, Schema Versions, and Key Terms

\hfill\break


\backmatter


\end{document}
